---
title: "IODS-final"
author: "Jaakko Keinänen"
date: "12/7/2017"
email: jaakko.keinanen@helsinki.fi
output: html_document
---
Data: Boston.

Question: what factors are associated with high crime rate? 

Plan: to test the question with logistic regression. The variable high_crime is true if the crime rate in the area is above the 3rd quartile of the average crime rate.

First looking at the distributions and correlations between high crime and other variables:
CRIM - per capita crime rate by town
ZN - proportion of residential land zoned for lots over 25,000 sq.ft.
INDUS - proportion of non-retail business acres per town.
CHAS - Charles River dummy variable (1 if tract bounds river; 0 otherwise)
NOX - nitric oxides concentration (parts per 10 million)
RM - average number of rooms per dwelling
AGE - proportion of owner-occupied units built prior to 1940
DIS - weighted distances to five Boston employment centres
RAD - index of accessibility to radial highways
TAX - full-value property-tax rate per $10,000
PTRATIO - pupil-teacher ratio by town
B - 1000(Bk - 0.63)^2 where Bk is the proportion of blacks by town
LSTAT - % lower status of the population
MEDV - Median value of owner-occupied homes in $1000's

```{r}
glimpse(Boston)
```

Let´s look at the correlations between crime and other variables.

```{r}
cor(Boston)

```

Visualization of correlations:
```{r}
ggpairs(Boston, mapping = aes(alpha=0.3), lower = list(combo = wrap("facethist", bins = 20)))
```


Hypothesis: High crime rate is predicted by lower status of population (lstat), low value of homes (medv), full-value property tax rate (tax) and pupil-teacher ratio (ptratio). These variables seem to be somewhat correlated with crime.

Logistic regression model to test the hypothesis:
```{r}
model <- glm(high_crime ~ lstat + medv + tax + ptratio, data = Boston, family="binomial")
```
Summary and coefficients of the model:
```{r}
summary(model)
coef(model)
```
Let´s compute odds ratios and confidence intervals for the model:
```{r}
library(dplyr)
OR <- coef(model) %>% exp
CI<- confint(model) %>% exp
cbind(OR, CI)
```
We can interpret from the odds ratios and CIs that lower status of population, full-value property-tax rate per $10,000 and pupil-teacher ratio by town are associated with higher crime rate. These variables have ORs and CI´s over 1.  
