---
title: "IODS-final"
author: "Jaakko Keinänen"
date: "12/15/2017"
email: jaakko.keinanen@helsinki.fi
output: html_document
---
Data: Boston.

## Wrangling:

Load Boston data from MASS library:

library(MASS)

summary(Boston)

str(Boston)

## Create new variable high_crime which is true if the crime is above 3rd quartile:

library(dplyr)

Boston <- mutate(Boston, high_crime= crim > 3.67708)

summary(Boston$high_crime)

write.csv(Boston, file="Bostoncrime.csv")

## End of wrangling.

Question: what factors are associated with high crime rate? 

Plan: to test the question with logistic regression. The variable high_crime is true if the crime rate in the area is above the 3rd quartile of the average crime rate.

First looking at the distributions and correlations between high crime and other variables:

CRIM - per capita crime rate by town

ZN - proportion of residential land zoned for lots over 25,000 sq.ft.

INDUS - proportion of non-retail business acres per town.

CHAS - Charles River dummy variable (1 if tract bounds river; 0 otherwise)

NOX - nitric oxides concentration (parts per 10 million)

RM - average number of rooms per dwelling

AGE - proportion of owner-occupied units built prior to 1940

DIS - weighted distances to five Boston employment centres

RAD - index of accessibility to radial highways

TAX - full-value property-tax rate per $10,000

PTRATIO - pupil-teacher ratio by town

B - 1000(Bk - 0.63)^2 where Bk is the proportion of blacks by town

LSTAT - % lower status of the population

MEDV - Median value of owner-occupied homes in $1000's

```{r}
Boston <- read.csv(file="~/Documents/Jaakon tutkimus/Open data kurssi s2017/GitHub/IODS-final/Bostoncrime.csv", header=T, sep = ",")
library(dplyr)
library(ggplot2)
library(GGally)
glimpse(Boston)
```

Let´s look at the correlations between crime and other variables.

```{r}
cor(Boston)

```

Visualization of correlations:
```{r fig.height=9, fig.width=9}
ggpairs(Boston, mapping = aes(alpha=0.3), lower = list(combo = wrap("facethist", bins = 20)))
```


Hypothesis: High crime rate is predicted by lower status of population (lstat), low value of homes (medv), full-value property tax rate (tax) and pupil-teacher ratio (ptratio). These variables seem to be somewhat correlated with crime.

```{r}

modelvars <- c("lstat", "medv", "tax", "ptratio", "high_crime")
modeldata <- select(Boston, one_of(modelvars))
summary(modeldata)
```


Logistic regression model to test the hypothesis:
```{r}
model <- glm(high_crime ~ lstat + medv + tax + ptratio, data = modeldata, family="binomial")
```
Summary and coefficients of the model:
```{r}
summary(model)
coef(model)
```
Let´s compute odds ratios and confidence intervals for the model:
```{r}
OR <- coef(model) %>% exp
CI<- confint(model) %>% exp
cbind(OR, CI)
```
We can interpret from the odds ratios and CIs that lower status of population, full-value property-tax rate per $10,000 and pupil-teacher ratio by town are associated with higher crime rate. These variables have ORs and CI´s over 1.  

Nex we´ll use the model to predict high crime. We will see whether using the model is better than guessing. We will add the predicted probabilities to the modeldata.

```{r}
probabilities <- predict(model, type = "response")
modeldata <- mutate(modeldata, probability = probabilities)
```
Using the probabilities to make predictions of high crime.
```{r}
modeldata <- mutate(modeldata, prediction = probability>0.5)
select(modeldata, lstat, medv, tax, ptratio, high_crime, probability, prediction) %>% tail(15)
```

Let´s tabulate the target variable versus the predictions:

```{r}
table(high_crime = modeldata$high_crime, prediction = modeldata$prediction)
table(high_crime = modeldata$high_crime, prediction = modeldata$prediction) %>% prop.table %>% addmargins()
```
From the first table we see that high crime is correctly predicted by the model in 368/379 false cases and in 126/127 of true cases. The model is much more accurate than just guessing.

The second table gives the proportions of FALSE and TRUE predictions and the target variable high_crime.

Plot high_crime vs probability:
```{r}
g <- ggplot(modeldata, aes(x = probability, y = high_crime, col=prediction))
g + geom_point()
```
We can see that the model is better in predicting correctly TRUE values than FALSE values.

Next we´ll construct a loss function which shows the incorrectly classified observations as loss and gives a measure of the accuracy of the model.
```{r}
loss_func <- function(class, prob) {
  n_wrong <- abs(class - prob) > 0.5
  mean(n_wrong)
}

loss_func(class = modeldata$high_crime, prob = modeldata$probability)
```
The output is the average prediction error.

